---
title: "Statsquatch: A Data Expedition into Bigfoot Sightings"
subtitle: "STA/ISS 313 - Project 1"
author: "The Tibbles"
format: html
editor: visual
---

## Abstract

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

The legend of Bigfoot has been a part of American lore for over a century. Over the years, there have been many recorded sighting of Bigfoot across the United States. The common geographic and weather factors between these sightings is the motivating question for this research. Using data from the Bigfoot Fied Researchers Organization website, we investigate the most common state and region for Bigfoot sightings and the weather conditions under which Bigfoot sightings are most common. Our findings hightlight trends in the geographical spread of Bigfoot sightings, including Bigfoot's recent reportings in Florida from Washington State, and report the most correlated weather conditions for sightings to occur, including 100% visibility conditions with humidity around 60%.

------------------------------------------------------------------------

## Introduction

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

The Tibbles choose the Bigfoot dataset from TidyTuesday. The data originates from a publicly available database on the Bigfoot Field Researchers Organization (BFRO) website, and the dataset created by Timothy Renner first became available on Data World in 2017. The dataset contains 5,021 rows, with each row representing a separate Bigfoot sighting, and 28 columns, with each column providing details on the sighting. The columns can generally be categorized as lengthier descriptions of the sighting observations, time and geographic details, and weather details.

The Tibbles opted to work with this dataset because of our shared predilection for the supernatural and the desire to know more about one of the most iconic creatures in the North American imagination. It contains a wealth of observations with plenty of categorical and numerical data well-suited for comprehensive analysis and interpretation.

## Question 1: How Does The Geographic Distribution Of Bigfoot Sightings Change Over Time?

### Introduction

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

Our first question looks at the geographic distribution of Bigfoot sightings, and how this distribution has changed over the decades. We use the latitude, longitude, and date variables to determine the precise location and decade the observation occurred.

We are interested in this because investigating changes in the geographic distribution of Bigfoot sightings over time could provide insights into patterns of human perception and belief in paranormal phenomena, as well as the impact of media and popular culture on these beliefs. Understanding these changes in the geographic distribution of sightings over time could also shed new light on this elusive creature.

### Approach

<<<<<<< HEAD
**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

=======
>>>>>>> 042be095ba22f4ac767b96d87f77a17a35fe1e1d
To address the first question, we utilized four different visualization techniques. The first method involved creating a scatter plot of Bigfoot sightings overlaid on a map of the mainland United States. This allowed us to visually observe clusters of sightings and areas of relative absence. The second method used a heat map of Bigfoot sightings overlayed on a map of the mainland United States. A heat map trades the precision of directly mapping all sightings for the advantage of a clear representation of density, which allowed for the identification of hot spots in regions with higher frequencies of Bigfoot sightings. We faceted both approaches by decade to facilitate comparisons across time.

The third and fourth methods employed stacked bar graphs to track the proportion of Bigfoot sightings recorded in each region of the country by decade. We used a stack bar graph to visualize changes in the distribution among regions over time. While determining the exact proportion for a region was slightly more difficult in a stacked bar graph, the density of data was much greater than for a conventional bar plot that might display regions in separate columns. We chose to graph proportions rather than nominal counts, as the focus of our research question was not the quantity of sightings, but rather the overall changes in the distribution as a whole. Additionally, some decades produced dramatically more sightings than others in the data set, making it difficult to observe trends in the less-populated decades of our data. Lastly, we used Census regions instead of states, since 50 individual colors would have been difficult to read (trading precision for clarity).

### Analysis

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

```{r question_1_SETUP, warning=FALSE, message=FALSE, echo=FALSE}
#install.packages("usmap")
#install.packages("mapdata")
library(knitr)
library(viridis)
library(tidyverse)
library(dplyr)
library(maps)
library(stringr)
library(mapdata)
bigfoot <- readr::read_csv('data/bigfoot.csv')
census <- readr::read_csv('data/CensusDivisions.csv')
bigfoot_xy <- bigfoot |> 
  select(latitude,
         longitude,
         date)
bigfoot_xy <- bigfoot_xy |> 
  filter(!is.na(latitude) & 
           !is.na(longitude) &
           !is.na(date))
bigfoot_xy <- bigfoot_xy |> 
  mutate(date = as.integer(substring(date,1,4)))
usa <- map_data('usa')
state <- map_data('state')
bigfoot_xy <- bigfoot_xy |> 
  mutate(alt_date = as.integer(substring(date,1,3))*10)
bigfoot_xy <- bigfoot_xy |> 
  mutate(decade = paste(substring(date,1,3),"0s",
                        sep = ""))
bigfoot_xy <- bigfoot_xy |> 
  mutate(decade=ifelse(date<1970,
                       "Pre-1970",
                       decade))
bigfoot_xy <- bigfoot_xy |> 
  mutate(decade=ifelse(date>2009,
                       "Post-2010",
                       decade))
bigfoot_xy <- bigfoot_xy |> 
  mutate(decade=factor(decade,
                       levels=c("Pre-1970",
                                "1970s",
                                "1980s",
                                "1990s",
                                "2000s",
                                "Post-2010")))

bigfoot_xy_mL <- bigfoot_xy |> filter(longitude > - 125 & 
                                         longitude < -65 & 
                                         latitude > 25 & 
                                         latitude < 50)
bigfoot_xy_mL_B <- bigfoot_xy_mL |> 
  group_by(decade) |> 
  summarise(longitude = median(longitude),
            latitude=median(latitude))
bigfoot_xy_mL_B <- bigfoot_xy_mL_B |>
  mutate(colour_dummy = "Median")
bigfoot_xy <- bigfoot_xy |>
  mutate(colour_dummy = "Sighting")
```

```{r question_1_MAP_PLOTS, fig.align="center", warning=FALSE, message=FALSE,fig.width=15, fig.height=12, fig.show='hold', echo=FALSE}
cols <- c("Sighting" = "darkgreen", 
          "Median" = "gold4")
ggplot()+
  geom_polygon(data=state, #State Boundary
               aes(x=long,
                   y=lat,
                   group=group),
               color="black",
               fill=NA,
               alpha=1)+
    geom_polygon(data=usa, #National Boundary
                 aes(x=long,
                     y=lat,
                     group=group),
                 color="black",
                 fill="white",
                 alpha=0.75)+
  geom_point(data=bigfoot_xy, #All Points
             mapping=aes(x=longitude,
                         y=latitude,
                         color=colour_dummy),
             shape=18,
             alpha=0.25)+
    geom_point(data = bigfoot_xy_mL_B, #Medians
               mapping=aes(x=longitude,
                           y=latitude,
                           color=colour_dummy),
               shape=18,
               size=5)+
  coord_fixed(ratio=1.3,
              xlim=c(-125,-65),
              ylim=c(25,50))+
    facet_wrap(~  decade,
               ncol=2)+
  labs(title="Bigfoot Sightings in the Mainland United States, 1869-present",
       axis.title=element_text(size=15),
        plot.title=element_text(size=25))+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      plot.title=element_text(size=20),
      panel.background=element_rect(fill="#EBEBEB"),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      legend.text=element_text(size=15),
      strip.text.x=element_text(size=15))+
      scale_colour_manual(values = cols,
                          name = element_blank())
rec_box <- data.frame(x=c(-155,-155,-55.5,-55.5,-155), 
                      y=c(10,60,60,10,10))
outline <- map("usa", 
               plot=FALSE)
outline <- data.frame(x=outline$x,
                      y=outline$y)
mask <- rbind(rec_box,
              outline)

ggplot()+
  geom_polygon(data=usa, #State Background
               aes(x=long,
                   y=lat,
                   group=group),
               color=NA,
               fill="white",
               alpha=1,
               linewidth=1)+
  geom_bin2d(data=bigfoot_xy, #Heatmap
             mapping=aes(x=longitude,
                         y=latitude),
             alpha=1,
             binwidth = c(2, 2))+
  geom_polygon(data=state, #State Boundary
               aes(x=long,
                   y=lat,
                   group=group),
               color="black",
               fill=NA,
               alpha=0.5)+
  geom_polygon(data=mask, #National Mask
               aes(x=x,y=y),
               color="#EBEBEB",
               fill="#EBEBEB",
               alpha=1)+
  geom_polygon(data=usa, #National Boundary
               aes(x=long,
                   y=lat,
                   group=group),
               color="black",
               fill=NA,
               alpha=1,
               linewidth=0.75)+
  coord_fixed(ratio=1.3,
    xlim=c(-125,-65),
              ylim=c(25,50))+
    facet_wrap(~  decade,
               ncol=2)+
  labs(title="Bigfoot Sightings in the Mainland United States, 1869-present",
       subtitle="Using 2° by 2° latitude/longitude grid")+
theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      plot.title=element_text(size=20),
      plot.subtitle=element_text(size=15),
      panel.background=element_blank(),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      strip.text.x=element_text(size=15))+
  scale_fill_gradientn(
    colours=c('#FFFC6B',
                       '#FF9700',
                       '#FF5B00',
                       '#E80000',
                       '#A80000',
                       '#790000'),
    limits=c(1, 100),
    name="Incidence")
```

```{r question_1_BARPLOTS, fig.align="center", warning=FALSE, message=FALSE, fig.show='hold',echo=FALSE}
cols_dv <- c("West South Central" = "firebrick1", 
          "East South Central" = "firebrick3",
          "South Atlantic" = "firebrick4",
          "West North Central" = "goldenrod1",
          "East North Central" = "goldenrod3",
          "Middle Atlantic" = "seagreen4",
          "New England" = "seagreen2",
          "Pacific" = "royalblue3",
          "Mountain" = "royalblue1")
cols_rg <- c("South" = "firebrick", 
          "Northeast" = "seagreen",
          "Midwest" = "goldenrod1",
          "West" = "royalblue")
bigfoot <- bigfoot |> filter(!is.na(date))
census <- readr::read_csv('data/CensusDivisions.csv')
census <- census |> rename("state" = "State",
                            "region" = "Region", 
                            "division" = "Division")
bigfoot <- left_join(bigfoot,
                     census, 
                     by="state")
bigfoot_sub <- bigfoot |> 
  select(state, 
         region, 
         division,
         date)
bigfoot_sub <- bigfoot_sub |> 
  mutate(year = as.integer(substring(date,1,4)))
bigfoot_sub <- bigfoot_sub |> 
  mutate(decade = paste(substring(year,1,3),"0s",
                        sep = ""))
bigfoot_sub <- bigfoot_sub |> 
  mutate(decade=ifelse(year<1970,
                       "Pre-1970",
                       decade))
bigfoot_sub <- bigfoot_sub |> 
  mutate(decade=ifelse(year>2009,
                       "Post-2010",
                       decade))
bigfoot_sub <- bigfoot_sub |>
  mutate(decade=factor(decade,
                       levels=c("Pre-1970",
                                "1970s",
                                "1980s",
                                "1990s",
                                "2000s",
                                "Post-2010")))
bigfoot_sub_rg <- bigfoot_sub |>
  group_by(region,
           decade)|>
  summarize(num = n())

bigfoot_sub_rg <- bigfoot_sub_rg |> 
  mutate(region=factor(region,
                       levels=c("Northeast",
                                "Midwest",
                                "South",
                                "West")))
ggplot(data = bigfoot_sub_rg,
       mapping = aes(fill=region, 
                     y=num, 
                     x=decade)) + 
    geom_bar(position="fill",
             stat="identity")+
  scale_fill_manual(values = cols_rg,
                          name = "U.S. Census\nRegion")+
  scale_y_continuous(labels = scales::percent)+
  labs(title="Geographic distribution of Bigfoot sightings, 1869-present")+
  xlab("Decade")+
  ylab("Proportion")+
  theme_bw()

bigfoot_sub_dv <- bigfoot_sub |>
  group_by(division,decade)|>
  summarize(num = n())

bigfoot_sub_dv <- bigfoot_sub_dv |> 
  mutate(division=factor(division,
                       levels=c("Middle Atlantic",
                                "New England",
                                "East North Central",
                                "West North Central",
                                "South Atlantic",
                                "East South Central",
                                "West South Central",
                                "Pacific",
                                "Mountain")))

state_mp <- state |> mutate(state=str_to_title(region))
state_mp <- left_join(state_mp, 
                      census,
                      by="state")
state_mp <- state_mp |> 
  mutate(division=factor(division,
                       levels=c("Middle Atlantic",
                                "New England",
                                "East North Central",
                                "West North Central",
                                "South Atlantic",
                                "East South Central",
                                "West South Central",
                                "Pacific",
                                "Mountain")))

state_mp <- state_mp |>
  mutate(region.y=factor(region.y,
                       levels=c("Northeast",
                                "Midwest",
                                "South",
                                "West")))

ggplot(data = bigfoot_sub_dv,
       mapping = aes(fill=division, 
                     y=num, 
                     x=decade)) + 
    geom_bar(position="fill", 
             stat="identity")+
  scale_fill_manual(values = cols_dv,
                          name = "U.S. Census\nDivision")+
  scale_y_continuous(labels = scales::percent)+
  labs(title="Geographic distribution of U.S. Bigfoot sightings, 1869-present")+
  xlab("Decade")+
  ylab("Proportion")+
  theme_bw()

```

```{r question_1_REFMAPS, fig.align="center", warning=FALSE, message=FALSE,fig.width=10, fig.height=5.2, fig.show='hold',echo=FALSE}
ggplot()+
  geom_polygon(data=state_mp,
               aes(x=long,
                   y=lat,
                   group=group,
                   fill=region.y),
               color="black",
               alpha=1)+
  scale_fill_manual(values = cols_rg,
                          name = "U.S. Census\nRegion")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      panel.background=element_rect(fill="white"),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.title=element_text(size=15),
      legend.title=element_text(size=12),
      legend.text=element_text(size=12),
      plot.caption=element_text(size=12),
      plot.margin = margin(t = 0, 
                             r = 0, 
                             b = 0, 
                             l = 0))+
      scale_colour_manual(values = cols_rg,
                          name = element_blank())+
  labs(caption="Hawaii and Alaska are in the West region",
       subtitle="Reference")+
  coord_fixed(ratio=1.3,
              xlim=c(-125,-65),
              ylim=c(25,50))

ggplot()+
  geom_polygon(data=state_mp,
               aes(x=long,
                   y=lat,
                   group=group,
                   fill=division),
               color="black",
               alpha=1)+
  scale_fill_manual(values = cols_dv,
                          name = "U.S. Census\nDivision")+
  theme(axis.line=element_blank(),
      axis.text.x=element_blank(),
      axis.text.y=element_blank(),
      axis.ticks=element_blank(),
      axis.title.x=element_blank(),
      axis.title.y=element_blank(),
      panel.background=element_rect(fill="white"),
      panel.grid.major=element_blank(),
      panel.grid.minor=element_blank(),
      plot.title=element_text(size=15),
      legend.title=element_text(size=12),
      legend.text=element_text(size=12),
      plot.caption=element_text(size=12),
      plot.margin = margin(t = 0, 
                             r = 0, 
                             b = 0, 
                             l = 0))+
      scale_colour_manual(values = cols_dv,
                          name = element_blank())+
  labs(caption="Hawaii and Alaska are in the Pacific division",
       subtitle="Reference")+
  coord_fixed(ratio=1.3,
              xlim=c(-125,-65),
              ylim=c(25,50))

```

### Discussion

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

Our figures demonstrate that the geographic distribution of Bigfoot sightings has undergone significant changes over time. A few findings are clear nationwide--- first, Bigfoot sightings have increased *everywhere* decade-over-decade. It is important to note, however, that earlier sightings may be less likely to appear in our data set. Beginning in the 2000s, Bigfoot was observed in areas where sightings had previously been absent, resulting in a wider geographic distribution, with sightings spanning the Dakotas, Plains States, and West Texas.

Another generalized finding--- Bigfoot sightings ebb and flow in specific areas over time. To name *just a few* examples:

-   During the 2000s, a range spanning from Eastern Texas to the Oklahoma/Arkansas border became a hot spot for Bigfoot sightings, despite having previously been a low-frequency area. The West South Central division, which includes Texas, Louisiana, Oklahoma, and Arkansas, doubled its proportion of national sightings from the 1990s to the 2000s, but this pattern dissipated after 2010.

-   Even though its lush wilderness would be perfect Bigfoot real estate, New England saw very few Bigfoot sightings until the 1990s. This has changed--- beginning in New York and moving eastward through Vermont, New Hampshire, all the way to Maine--- Bigfoot has arrived, and this is visible in the heat-map. The New England census division was essentially barren of any sightings in the 1980s, but post-2010, it now contributes about 5% of the national total, outpacing its share of area.

-   During the 2000s, Bigfoot sightings increased dramatically in the vicinity of Lake Tahoe and the Eldorado National Forest in California (identifiable both in the heat-map and scatter-plot). The grid-space containing this region actually set a record in the 2000s for the highest incidence outside of Washington State. Like the Texas hot spot, after 2010, this pattern completely disappeared. California is in the same region and division as Washington, limiting the conclusions we can draw from our bar plot that apply specifically to this pattern, but their (Pacific) division did see its proportion of sightings plummet post-2010, buttressing our conclusion from the heat-map.

***(should this be another bullet point???)***Lastly, Bigfoot is moving east! After 2010, the median Bigfoot sighting jumped the Mississippi River into Illinois for the first time ever in our data. The heat-map explains why: downstate Illinois, northeast Ohio, and the Hudson River Valley have all become hot spots for Bigfoot sightings (the Ohio cluster actually dates all the way back to the 1970s). The West region of the U.S. --- spanning from Honolulu to Denver--- contributed 40% of the nation's sightings in the 1990s, but this has fallen to only a quarter post-2010, actually disproportionately *less* than its area would warrant.

What this data, taken as a whole, tell us, is that the geographic distribution of sightings is highly variable. We venture two explanations: first, that Bigfoot's tastes themselves are changing. Like all creatures, Bigfoot may have gotten bored and spread his wings- maybe for a yeti, Lake Tahoe was the place to be in 2007. Alternatively, we argue, Bigfoot sightings are a highly social event--- people think they see Bigfoot and tell their family, friends, and even the media--- making it much more likely that the individuals exposed to that information will earnestly (or for attention) claim to sight Bigfoot later. If one sighting can 'trigger' another, it would be logical for them to appear in geographical clusters like we observed forming and dissipating. While we cannot conclusively resolve the matter of Bigfoot's existence, we hypothesize that the latter explanation is more likely.

## Question 2: Weather Conditions and their Effects on Bigfoot Sightings

### Introduction

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

Question 2 investigates the relationship between weather conditions and the number of Bigfoot sightings. The analysis focuses on seven weather variables, including dew point, humidity, temperature, cloud cover, UV index, visibility, wind bearing, and wind speed. By examining changes in these variables over time and their association with the number of Bigfoot sightings, we aim to gain a better understanding of the conditions that may be associated with Bigfoot sightings.

We are interested in investigating the relationship between weather and Bigfoot sightings because it could contribute to a better understanding of the factors that influence human perceptions of the natural world. In addition, it could lead to new discoveries or hypotheses about the nature of Bigfoot.

### Approach

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

### Analysis

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

(2-3 code blocks, 2 figures, text/code comments as needed): In this section, provide the code that generates your plots. Use scale functions to provide nice axis labels and guides. You are welcome to use theme functions to customize the appearance of your plot, but you are not required to do so. All plots must be made with **ggplot2**. Do not use base R or lattice plotting functions.

```{r question-two, message=FALSE, warning=FALSE, echo=FALSE, fig.show="hold", fig.width=10, fig.height=5}
library(dplyr)
library(tidyverse)
library(ggridges)
library(knitr)
bigfoot <- readr::read_csv('data/bigfoot.csv')
bigfoot <- bigfoot |>
  select(dew_point,
         humidity,
         temperature_mid,
         cloud_cover,
         uv_index,
         visibility,
         wind_bearing,
         wind_speed) |>
  mutate(y_dummy = 1)

cols_qt <- c("1" = "#519A68", 
          "2" = "#51837A",
          "3" = "#517387",
          "4" = "#515E97")




# Dew Point
ggplot(data=bigfoot |>
         filter(!is.na(dew_point)),
       mapping = aes(x = dew_point,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Dew Point (°F)")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                      values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by dew point",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$dew_point,na.rm=TRUE) / 
                               mean(bigfoot$dew_point,
                                    na.rm=TRUE),
                             digits = 4)))

# Dew Point
ggplot(data=bigfoot |>
         filter(!is.na(humidity)),
       mapping = aes(x = humidity,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Humidity")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  scale_x_continuous(labels = scales::percent)+
  labs(title="Distribution of Bigfoot sightings by humidity",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$humidity,na.rm=TRUE) / 
                               mean(bigfoot$humidity,
                                    na.rm=TRUE),
                             digits = 4)))

# Temperature Mid
ggplot(data=bigfoot |>
         filter(!is.na(temperature_mid)),
       mapping = aes(x = temperature_mid,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Average Temperature for Day of Sighting (°F)")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by temperature",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$temperature_mid,na.rm=TRUE) / 
                               mean(bigfoot$temperature_mid,
                                    na.rm=TRUE),
                             digits = 4)))

# Cloud Cover
ggplot(data=bigfoot |>
         filter(!is.na(cloud_cover)),
       mapping = aes(x = cloud_cover,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Cloud Cover")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  scale_x_continuous(labels = scales::percent)+
  labs(title="Distribution of Bigfoot sightings by cloud cover",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$cloud_cover,na.rm=TRUE) / 
                               mean(bigfoot$cloud_cover, 
                                    na.rm=TRUE),
                             digits = 4)))

# UV
ggplot(data=bigfoot |>
         filter(!is.na(uv_index)),
       mapping = aes(x = uv_index,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("U.V. Index")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by U.V. Index",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$uv_index,na.rm=TRUE) /
                               mean(bigfoot$uv_index,
                                    na.rm=TRUE),
                             digits = 4)))

# Visibility
ggplot(data=bigfoot |>
         filter(!is.na(visibility)),
       mapping = aes(x = visibility,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Visibility (mi)")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by visibility",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$visibility,na.rm=TRUE) /
                               mean(bigfoot$visibility,
                                    na.rm=TRUE),
                             digits = 4)))

# Wind Bearing
ggplot(data=bigfoot |>
         filter(!is.na(wind_bearing)),
       mapping = aes(x = wind_bearing,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Wind Bearing (°)")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by wind bearing",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$wind_bearing,na.rm=TRUE) / 
                               mean(bigfoot$wind_bearing,
                                    na.rm=TRUE),
                             digits = 4)))

# Wind Speed
ggplot(data=bigfoot |>
         filter(!is.na(wind_speed)),
       mapping = aes(x = wind_speed,
                     y = y_dummy,
                     fill = factor(stat(quantile))))+ 
  stat_density_ridges(geom = "density_ridges_gradient",
                      quantile_lines = TRUE, 
                      quantiles = 4, 
                      calc_ecdf = TRUE)+
  xlab("Wind Speed (mph)")+
  theme(axis.text.y=element_blank(),
          axis.ticks.y=element_blank(),
        axis.title.y=element_blank(),
        aspect.ratio=.5)+
  scale_fill_manual(name="Quartile",
                    values = cols_qt)+
  labs(title="Distribution of Bigfoot sightings by wind speed",
       subtitle=paste("Relative standard deviation (RSD):",
                      format(sd(bigfoot$wind_speed,na.rm=TRUE) / 
                               mean(bigfoot$wind_speed,
                                    na.rm=TRUE),
                             digits = 4)))

Variable <- c("Dew Point", 
          "Humidity", 
          "Avg. Temperature", 
          "Cloud Cover", 
          "U.V. Index",
          "Visibility",
          "Wind Bearing",
          "Wind Speed")
Dispersion <- c((sd(bigfoot$dew_point,
                    na.rm=TRUE) /
                   mean(bigfoot$dew_point,
                        na.rm=TRUE)),
         (sd(bigfoot$humidity,
             na.rm=TRUE) / 
            mean(bigfoot$humidity,
                 na.rm=TRUE)), 
         (sd(bigfoot$temperature_mid,
             na.rm=TRUE) / 
            mean(bigfoot$temperature_mid,
                 na.rm=TRUE)),
         (sd(bigfoot$cloud_cover,
             na.rm=TRUE) / 
            mean(bigfoot$cloud_cover,
                 na.rm=TRUE)),
         (sd(bigfoot$uv_index
             ,na.rm=TRUE) / 
            mean(bigfoot$uv_index,
                 na.rm=TRUE)),
         (sd(bigfoot$visibility,
             na.rm=TRUE) / 
            mean(bigfoot$visibility,
                 na.rm=TRUE)),
         (sd(bigfoot$wind_bearing,
             na.rm=TRUE) / 
            mean(bigfoot$wind_bearing,
                 na.rm=TRUE)),
         (sd(bigfoot$wind_speed,
             na.rm=TRUE) / 
            mean(bigfoot$wind_speed,
                 na.rm=TRUE)))

df <- data.frame(Variable, 
                 Dispersion)

df <- df |>
  arrange(Dispersion)

knitr::kable(df, 
             col.names=c("Variable",
                         "Relative Standard Deviation"),
             caption="Dispersion of Weather Variables",
             digits=4)


ggplot(bigfoot, aes(visibility, humidity))+
  geom_bin_2d()+
  scale_fill_continuous(name="Count")+
  labs(title="Humidity and Visibility Conditions during Bigfoot Sightings",
       subtitle="Variables selected on the basis of low dispersion")+
  xlab("Visibility (mi)")+
  ylab("Humidity")+
  scale_y_continuous(labels=scales::percent)

ggplot(bigfoot, aes(visibility, temperature_mid))+
  geom_bin_2d()+
  scale_fill_continuous(name="Count")+
  labs(title="Temperature and Visibility Conditions during Bigfoot Sightings",
       subtitle="Variables selected on the basis of low dispersion")+
  xlab("Visibility (mi)")+
  ylab("Average Temperature for Day of Sighting (°F)")

ggplot(bigfoot, aes(temperature_mid, humidity))+
  geom_bin_2d()+
  scale_fill_continuous(name="Count")+
  labs(title="Humidity and Temperature Conditions during Bigfoot Sightings",
       subtitle="Variables selected on the basis of low dispersion")+
  xlab("Average Temperature for Day of Sighting (°F)")+
  ylab("Humidity")+
  scale_y_continuous(labels=scales::percent)

```

[***!!! NOT COMPLETE !!!***]{.underline}

### Discussion

**READ AND REVIEWED BY ANIK**

**READ AND REVIEWED BY ARYAMAN**

**READ AND REVIEWED BY MATT**

**READ AND REVIEWED BY NHU**

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.

[***!!! NOT COMPLETE !!!***]{.underline}
