---
title: "Statsquatch: A Data Expedition into Bigfoot Sightings"
subtitle: "STA/ISS 313 - Project 1"
author: "The Tibbles"
format: html
editor: visual
---

## Abstract

The legendary cryptid Bigfoot (also commonly known as .Sasquatch, Yeti, or Yowie) has been a vibrant part of American lore for over a century. The data used for this project was adapted by Timothy Renner from the Bigfoot Field Researchers Organization digital database, and it provides details on the thousands of sighting in the United States recorded over the years -- with the first observation dating all the way back to 1869. Given the focus of data on geographic and atmospheric variables surrounding the sightings, our motivating questions revolve around uncovering the relationship between these variables and observations of Bigfoot. We first investigated the most common state and region for Bigfoot sightings, finding that sightings in the West region are consistently most common over time -- specifically in the Pacific states. Next, we investigated the relationship between precipitation conditions on the day of Bigfoot sightings and the number of sightings and found that [**INSERT DISCUSSION FROM Q2**]{.underline}. Through these questions, this project sufficiently analyzes and highlights trends in some of the geographic and weather conditions most commonly associated with Bigfoot sightings.

------------------------------------------------------------------------

## Introduction

The Tibbles choose the Bigfoot dataset from TidyTuesday. The data originates from a digital database publically available on the Bigfoot Field Researchers Organization (BFRO) website, and the dataset itself, which first became available on Data World in 2017, was created by Timothy Renner. The dataset contains 5,021 rows, with each row representing a separate Bigfoot sighting, and 28 columns, with each column providing details on the sighting. The columns can generally be categorized as either lengthier verbal descriptions of the sighting observations or a combination of numerical and categorical variables that describe details regarding time, geographic, and weather conditions.

The Tibbles opted to work with this dataset because of our shared predilection for the supernatural and the desire to know more about one of the most iconic creatures in the North American imagination. It contains a wealth of observations with a variety of associated categorical and numerical data well-suited for comprehensive analysis and interpretation.

## Question 1: How does the geographic distribution of Bigfoot sightings change over time?

### Introduction

Our first question investigates the geographic distribution of Bigfoot sightings and how this distribution has changed over time, using the latitude, longitude, and date variables to determine the precise location and decade the observation occurred. Analyzing the distribution of sightings and furthermore changes over times could provide insights on patterns of the public's relationship with paranormal phenomena, as well as the impact of media and popular culture on these beliefs. Each region and each generation in turn has different social and cultural beliefs, formed by the overarching history of the area and the events of the decade. Understanding these changes in the geographic distribution of sightings over time could also shed new light on this elusive creature.

### Approach

To address the first question, we utilized three different visualization techniques. The first method involved creating a scatter plot of Bigfoot sightings overlaid on a map of the mainland United States. This allowed us to visually observe clusters of sightings and areas of relative absence. The second method used a heat map of Bigfoot sightings overlayed on a map of the mainland United States. A heat map trades the precision of directly mapping all sightings for the advantage of a clear representation of density, which allowed for the identification of hot spots in regions with higher frequencies of Bigfoot sightings. We faceted both approaches by decade to facilitate comparisons across time.

The third method employed a stacked bar graph to track the proportion of Bigfoot sightings recorded in each division of the country by decade. We used a stack bar graph to visualize changes in the distribution among divisions over time. While determining the exact proportion for a division was slightly more difficult in a stacked bar graph, the density of data was much greater than for a conventional bar plot that might display divisions in separate columns. We chose to graph proportions rather than nominal counts, as the focus of our research question was not the quantity of sightings, but rather the overall changes in the distribution as a whole. Additionally, some decades produced dramatically more sightings than others in the data set, making it difficult to observe trends in the less-populated decades of our data.

### Analysis

```{r}
#| label: question-1-MAP_PLOTS
#| fig.align: "center"
#| warning: FALSE
#| message: FALSE
#| fig.width: 15
#| fig.height: 12
#| fig.show: 'hold'
#| echo: FALSE

#SET-UP

#install.packages("usmap")
#install.packages("mapdata")

library(knitr)
library(viridis)
library(tidyverse)
library(dplyr)
library(maps)
library(stringr)
library(mapdata)

bigfoot <- readr::read_csv('data/bigfoot.csv')
census <- readr::read_csv('data/CensusDivisions.csv')
usa <- map_data('usa')
state <- map_data('state')


#DATA-WRANGLING

bigfoot_xy <- bigfoot |>
  select(
    latitude,
    longitude,
    date
  ) |>
  filter(!is.na(latitude) &
    !is.na(longitude) &
    !is.na(date)) |>
  mutate(date = as.integer(substring(date, 1, 4))) |>
  mutate(alt_date = as.integer(substring(date, 1, 3)) * 10) |>
  mutate(decade = paste(substring(date, 1, 3), "0s",
    sep = ""
  )) |>
  mutate(decade = ifelse(date < 1970,
    "Pre-1970",
    decade
  )) |>
  mutate(decade = ifelse(date > 2009,
    "Post-2010",
    decade
  )) |>
  mutate(decade = factor(decade,
    levels = c(
      "Pre-1970",
      "1970s",
      "1980s",
      "1990s",
      "2000s",
      "Post-2010"
    )
  )) |>
  mutate(colour_dummy = "Sighting")

bigfoot_xy_mL <- bigfoot_xy |>
  filter(
    longitude > -125 &
      longitude < -65 &
      latitude > 25 &
      latitude < 50
  )

bigfoot_xy_mL_B <- bigfoot_xy_mL |>
  group_by(decade) |>
  summarise(
    longitude = median(longitude),
    latitude = median(latitude)
  ) |>
  mutate(colour_dummy = "Median")


#SCATTERPLOT MAP

cols <- c(
  "Sighting" = "darkgreen",
  "Median" = "gold4"
)

ggplot() +
  geom_polygon(
    data = state, # State Boundary
    aes(
      x = long,
      y = lat,
      group = group
    ),
    color = "black",
    fill = NA,
    alpha = 1
  ) +
  geom_polygon(
    data = usa, # National Boundary
    aes(
      x = long,
      y = lat,
      group = group
    ),
    color = "black",
    fill = "white",
    alpha = 0.75
  ) +
  geom_point(
    data = bigfoot_xy, # All Points
    mapping = aes(
      x = longitude,
      y = latitude,
      color = colour_dummy
    ),
    shape = 18,
    alpha = 0.25
  ) +
  geom_point(
    data = bigfoot_xy_mL_B, # Medians
    mapping = aes(
      x = longitude,
      y = latitude,
      color = colour_dummy
    ),
    shape = 18,
    size = 5
  ) +
  coord_fixed(
    ratio = 1.3,
    xlim = c(-125, -65),
    ylim = c(25, 50)
  ) +
  facet_wrap(~decade,
    ncol = 2
  ) +
  labs(
    title = "Bigfoot Sightings in the Mainland United States, 1869-present",
    axis.title = element_text(size = 15),
    plot.title = element_text(size = 25)
  ) +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = 20),
    panel.background = element_rect(fill = "#EBEBEB"),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    legend.text = element_text(size = 15),
    strip.text.x = element_text(size = 15)
  ) +
  scale_colour_manual(
    values = cols,
    name = element_blank()
  )
rec_box <- data.frame(
  x = c(-155, -155, -55.5, -55.5, -155),
  y = c(10, 60, 60, 10, 10)
)
outline <- map("usa",
  plot = FALSE
)
outline <- data.frame(
  x = outline$x,
  y = outline$y
)
mask <- rbind(
  rec_box,
  outline
)


#HEAT MAP

ggplot() +
  geom_polygon(
    data = usa, # State Background
    aes(
      x = long,
      y = lat,
      group = group
    ),
    color = NA,
    fill = "white",
    alpha = 1,
    linewidth = 1
  ) +
  geom_bin2d(
    data = bigfoot_xy, # Heatmap
    mapping = aes(
      x = longitude,
      y = latitude
    ),
    alpha = 1,
    binwidth = c(2, 2)
  ) +
  geom_polygon(
    data = state, # State Boundary
    aes(
      x = long,
      y = lat,
      group = group
    ),
    color = "black",
    fill = NA,
    alpha = 0.5
  ) +
  geom_polygon(
    data = mask, # National Mask
    aes(x = x, y = y),
    color = "#EBEBEB",
    fill = "#EBEBEB",
    alpha = 1
  ) +
  geom_polygon(
    data = usa, # National Boundary
    aes(
      x = long,
      y = lat,
      group = group
    ),
    color = "black",
    fill = NA,
    alpha = 1,
    linewidth = 0.75
  ) +
  coord_fixed(
    ratio = 1.3,
    xlim = c(-125, -65),
    ylim = c(25, 50)
  ) +
  facet_wrap(~decade,
    ncol = 2
  ) +
  labs(
    title = "Bigfoot Sightings in the Mainland United States, 1869-present",
    subtitle = "Using 2° by 2° latitude/longitude grid"
  ) +
  theme(
    axis.line = element_blank(),
    axis.text.x = element_blank(),
    axis.text.y = element_blank(),
    axis.ticks = element_blank(),
    axis.title.x = element_blank(),
    axis.title.y = element_blank(),
    plot.title = element_text(size = 20),
    plot.subtitle = element_text(size = 15),
    panel.background = element_blank(),
    panel.grid.major = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text.x = element_text(size = 15)
  ) +
  scale_fill_gradientn(
    colours = c(
      "#FFFC6B",
      "#FF9700",
      "#FF5B00",
      "#E80000",
      "#A80000",
      "#790000"
    ),
    limits = c(1, 100),
    name = "Incidence"
  )
```

```{r}
#| label: question-1-BARPLOT
#| fig.align: "center"
#| warning: FALSE
#| message: FALSE
#| fig.show: 'hold'
#| echo: FALSE

#DATA-WRANGLING

cols_dv <- c(
  "West South Central" = "firebrick1",
  "East South Central" = "firebrick3",
  "South Atlantic" = "firebrick4",
  "West North Central" = "goldenrod1",
  "East North Central" = "goldenrod3",
  "Middle Atlantic" = "seagreen4",
  "New England" = "seagreen2",
  "Pacific" = "royalblue3",
  "Mountain" = "royalblue1"
)

bigfoot <- bigfoot |> filter(!is.na(date))
census <- census |> rename(
  "state" = "State",
  "region" = "Region",
  "division" = "Division"
)

bigfoot <- left_join(bigfoot,
  census,
  by = "state"
)

bigfoot_sub <- bigfoot |>
  select(
    state,
    region,
    division,
    date
  ) |>
  mutate(year = as.integer(substring(date, 1, 4))) |>
  mutate(decade = paste(substring(year, 1, 3), "0s",
    sep = ""
  )) |>
  mutate(decade = ifelse(year < 1970,
    "Pre-1970",
    decade
  ))|>
  mutate(decade = ifelse(year > 2009,
    "Post-2010",
    decade
  )) |>
  mutate(decade = factor(decade,
    levels = c(
      "Pre-1970",
      "1970s",
      "1980s",
      "1990s",
      "2000s",
      "Post-2010"
    )
  ))

bigfoot_sub_dv <- bigfoot_sub |>
  group_by(division, decade) |>
  summarize(num = n()) |>
  mutate(division = factor(division,
    levels = c(
      "Middle Atlantic",
      "New England",
      "East North Central",
      "West North Central",
      "South Atlantic",
      "East South Central",
      "West South Central",
      "Pacific",
      "Mountain"
    )
  ))


#BARPLOT

ggplot(
  data = bigfoot_sub_dv,
  mapping = aes(
    fill = division,
    y = num,
    x = decade
  )
) +
  geom_bar(
    position = "fill",
    stat = "identity"
  ) +
  scale_fill_manual(
    values = cols_dv,
    name = "U.S. Census\nDivision"
  ) +
  scale_y_continuous(labels = scales::percent) +
  labs(title = "Geographic distribution of U.S. Bigfoot sightings, 1869-present") +
  xlab("Decade") +
  ylab("Proportion") +
  theme_bw()
```

### Discussion

Our figures demonstrate that the geographic distribution of Bigfoot sightings has undergone significant changes over time. A few findings are clear nationwide--- first, Bigfoot sightings have increased *everywhere* decade-over-decade. It is important to note, however, that earlier sightings may be less likely to appear in our data set. Beginning in the 2000s, Bigfoot was observed in areas where sightings had previously been absent, resulting in a wider geographic distribution, with sightings spanning the Dakotas, Plains States, and West Texas.

Another generalized finding--- Bigfoot sightings ebb and flow in specific areas over time. To name *just a few* examples:

-   During the 2000s, a range spanning from Eastern Texas to the Oklahoma/Arkansas border became a hot spot for Bigfoot sightings, despite having previously been a low-frequency area. The West South Central division, which includes Texas, Louisiana, Oklahoma, and Arkansas, doubled its proportion of national sightings from the 1990s to the 2000s, but this pattern dissipated after 2010.

-   Even though its lush wilderness would be perfect Bigfoot real estate, New England saw very few Bigfoot sightings until the 1990s. This has changed--- beginning in New York and moving eastward through Vermont, New Hampshire, all the way to Maine--- Bigfoot has arrived, and this is visible in the heat-map. The New England census division was essentially barren of any sightings in the 1980s, but post-2010, it now contributes about 5% of the national total, outpacing its share of area.

-   During the 2000s, Bigfoot sightings increased dramatically in the vicinity of Lake Tahoe and the Eldorado National Forest in California (identifiable both in the heat-map and scatter-plot). The grid-space containing this region actually set a record in the 2000s for the highest incidence outside of Washington State. Like the Texas hot spot, after 2010, this pattern completely disappeared. California is in the same region and division as Washington, limiting the conclusions we can draw from our bar plot that apply specifically to this pattern, but their (Pacific) division did see its proportion of sightings plummet post-2010, buttressing our conclusion from the heat-map.

Lastly, Bigfoot is moving east! After 2010, the median Bigfoot sighting jumped the Mississippi River into Illinois for the first time ever in our data. The heat-map explains why: downstate Illinois, northeast Ohio, and the Hudson River Valley have all become hot spots for Bigfoot sightings (the Ohio cluster actually dates all the way back to the 1970s). The West region of the U.S. --- spanning from Honolulu to Denver--- contributed 40% of the nation's sightings in the 1990s, but this has fallen to only a quarter post-2010, actually disproportionately *less* than its area would warrant.

What this data, taken as a whole, tell us, is that the geographic distribution of sightings is highly variable. We venture two explanations: first, that Bigfoot's tastes themselves are changing. Like all creatures, Bigfoot may have gotten bored and spread his wings- maybe for a yeti, Lake Tahoe was the place to be in 2007. Alternatively, we argue, Bigfoot sightings are a highly social event--- people think they see Bigfoot and tell their family, friends, and even the media--- making it much more likely that the individuals exposed to that information will earnestly (or for attention) claim to sight Bigfoot later. If one sighting can 'trigger' another, it would be logical for them to appear in geographical clusters like we observed forming and dissipating. While we cannot conclusively resolve the matter of Bigfoot's existence, we hypothesize that the latter explanation is more likely.

## Question 2: What precipitation conditions are most commonly associated with Bigfoot sightings?

### Introduction

Our second question investigates the relationship between precipitation and the number of Bigfoot sightings. The analysis focuses on two of the three precipitation-related variables: precip_probability and precip_type. By examining the relationship between the precipitation conditions and the number of Bigfoot sightings, we attempt to derive and explain an association between the conditions and Bigfoot sightings. The legends surrounding precipitation weather events eclipse those surrounding Bigfoot. Thus, precipitation data and may be able to explain the sightings from in two ways: these variables can contribute to a better understanding of an important physical weather factor that can influence human perception of the cryptic world, and/or it can lead to new discoveries and hypotheses about the nature of Bigfoot and his existence.

### Approach

Note: The proposal originally outlined a plan to evaluate the distribution of a variety of weather variables and compare results. However, The Tibbles ultimately decided to focus on a single weather conditions which had multiple variables to describe it: precipitation, which has variables describing the precipitation type, intensity, and probability. Other than temperature, it is the only weather condition with multiple variables, so precipitation lends itself to more thorough analysis --- compounded by the ideas described in the introduction. After creating some preliminary visualizations, The Tibbles also decided that it would be most prudent to only analyze type and probability because the values in the intensity column were extremely heavily concentrated at or near 0.

To address the second question, we first began with boxplot describing the median probability of precipitation, with a separate boxplot for each season. Little is understood about how the probability of precipitation is measured and meant to be understood; it is a calculation combining how likely the precipitation event is and what percentage of the area of the forecast the predicted event is predicted to cover. Given that metereologists typically have quite a high level of precipitation prediction accuracy, The Tibbles assume that the probability value is more relevant to the percentage of area coverage. Because all the weather data is generalized to the day and city/locality of the sighting and not to the exact time and site of the sighting, it is certainly important in order to evaluate the relevance of precipitation to Bigfoot sightings to first evaluate the relevance of that precipitation to the actual location of the sighting. The boxplot is the best plot for this goal because it allows one to easily compare and evaluate the median probabilities for each season (labeled), while still providing important context about the range of probabilities that exist.

The second plot is a barplot comparing the number of sightings in relation to whether there was precipitation or not, and there is also the distinction of what type of precipitation was present (rain or snow). It is faceted by season to allow for the accounting of the profound impact that season can have on the presence and type of precipitation in our analysis. The barplot allows for multiple easy comparisons, especially across facets: it allows for the comparison between number of sightings with and without precipitation, between precipitation within the presence of precipitation, and the previous two comparisons but between seasons. Given that in the United States, the majority of days throughout the year do not have precipitation, even a small difference between the number of sightings with and without precipitation can be telling. All these factors will be important in evaluating the existence and extent of an association between precipitation and a Bigfoot sighting.

### Analysis

```{r}
#| label: question-2-PROBABILITY_PLOT
#| warning: FALSE
#| message: FALSE
#| echo: FALSE
#| fig.show: "hold"

#SET-UP

library(scales)


#DATA WRANGLING

precip_probability_median <- bigfoot |>
  filter(!is.na(precip_probability)) |>
  filter(precip_probability > 0) |>
  filter(season != "Unknown") |>
  group_by(season) |>
  summarize(median = median(precip_probability)) |>
  mutate(median_percent = percent(median))


#PRECIPITATION PROBABILITY PLOT

bigfoot |>
  filter(!is.na(precip_probability)) |>
  filter(precip_probability > 0) |>
  filter(season != "Unknown") |>
  ggplot(aes(x = precip_probability, y = season)) +
  geom_boxplot() +
  geom_text(
    data = precip_probability_median,
    aes(
      x = median, y = season, label = median_percent
    ), size = 3, hjust = 1.5
  ) +
  scale_x_continuous(labels = c("0%", "25%", "50%", "75%", "100%")) +
  scale_y_discrete(labels = c("Fall", "Summer", "Spring", "Winter")) +
  coord_cartesian(clip = "off") +
  labs(
    title = "Precipitation Probability during Bigfoot Sightings by Season",
    x = "Probability of Precipitation",
    y = "Season"
  ) +
  theme_minimal()
```

```{r}
#| label: question-2-TYPE_PLOT
#| message: FALSE
#| warning: FALSE
#| echo: FALSE
#| fig.show: "hold"

#SET-UP

library(ggridges)


#DATA WRANGLING

cols_rg <- c("None" = "goldenrod1", 
          "Rain" = "royalblue",
          "Snow" = "lightblue")

bigfoot_precip <- bigfoot |>
  mutate(
    precip_type = ifelse(
      precip_type == "rain", "Rain", ifelse(precip_type == "snow", "Snow", precip_type)
    )
  ) |>
  mutate(
    precip_type =
      ifelse(is.na(precip_type),
        ifelse(!is.na(temperature_mid), "None", precip_type), precip_type
      )
  ) |>
  select(
    precip_type, season, region
  ) |>
  filter(season != "Unknown" & !is.na(precip_type)) |>
  mutate(precipitation = ifelse(precip_type == "None", "No", "Yes"))


#PRECIPITATION TYPE BARPLOT

bigfoot_precip |>
  ggplot() +
  geom_bar(mapping = aes(y = precipitation, fill = precip_type)) +
  facet_wrap(
    ~ factor(
      season,
      levels = c("Fall", "Summer", "Spring", "Winter")
    ),
    ncol = 1
  ) +
  scale_fill_manual(
    values = cols_rg,
    name = "Precipitation Type"
  ) +
  labs(
    title = "Bigfoot Sightings by Precipitation: Singin' in the Rain",
    x = "Bigfoot Sightings",
    y = "Precipitation?"
  ) +
  theme_minimal()
```

### Discussion

(1-3 paragraphs): In the Discussion section, interpret the results of your analysis. Identify any trends revealed (or not revealed) by the plots. Speculate about why the data looks the way it does.
