---
title: "Statsquatch: A Data Analysis of Bigfoot Sightings"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
suppressWarnings(library(tidyverse))
library(knitr)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

bigfoot <- readr::read_csv('data/bigfoot.csv')

glimpse(bigfoot)
```

The Tibbles chose the Bigfoot dataset from TidyTuesday. The data originates from a publicly available database on the Bigfoot Field Researchers Organization (BFRO) website, and the dataset created by Timothy Renner first became available on Data World in 2017. The dataset contains 5,021 rows, with each row representing a separate Bigfoot sighting, and 28 columns, with each column providing details on the sighting. The columns can generally be categorized as lengthier descriptions of the sighting observations, time and geographic details, and weather details.

The Tibbles opted to work with this dataset because of our shared predilection for the supernatural and the potential this information holds to prove once and for all the existence (or lack thereof) of one of the most iconic creatures in the North American imagination. It contains a wealth of observations with plenty of categorical and numerical data well-suited for comprehensive analysis and interpretation.

## Questions

1.  How has the geographic distribution of Bigfoot sightings changed over time?
2.  Are certain weather conditions more commonly associated with Bigfoot sightings than others?

## Analysis plan

-   **Question One: How has the geographic distribution of Bigfoot sightings changed over time?**
    -   Create new 'year' variable from existing 'date' variable using the mutate() function

    -   **Figure 1:** Map observations as points by 'longitude' and 'latitude' variables then color by 'year' using a gradient color scale (oldest values shaded darker, newest values shaded brighter, e.g.). If the data points stack to the extent of obscuring the trend, we may (a) explore using transparency to improve visibility, or (b) use a heat-map instead, plotting point density faceted by decade.

    -   **Figure 2:** Use a bar plot to count the number of sighting in each state, facet this graph by year. This visualization will show which states have the most Bigfoot sightings and how that changes over time.

*Further Plan for Evaluation*

For Figure 1, we will use the map plot to perform a visual observation of any possible trends in the geographic distribution. For example, a cluster of specific points during the 1980s from one area might be evidence of an unusual pattern (perhaps a national park, e.g.) that would be worth further observation. The Professor notes -- correctly -- that it may be difficult to draw formal conclusions from such a plot; however, we believe that when provided with geographic data, good practice in a comprehensive overview would include taking a look *at* that data in a raw format, and mapping it, even if it proves useless, before we apply arbitrary strata (like states). We have also adopted a few contingency options in case the first draft of the map is unusable.

-   **Question Two: Are certain weather conditions more commonly associated with Bigfoot sightings than others?**
    -   Create a histogram and ridgeline plot (subject to change) illustrating the distribution of Bigfoot sightings with respect to 'cloud_cover', 'moon_phase', 'precip_intensity', 'precip_type', and 'visibility'

    -   Analyze plot for patterns in distribution

    -   Create additional plots on weather variables if necessary. Weather variables include:

        -   temperature_low/mid/high

        -   dew_point

        -   humidity
