---
title: "Statsquatch: A Data Analysis of Bigfoot Sightings"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false

library(tidyverse)
```

# Dataset

```{r}
#| label: load-dataset
#| message: false

bigfoot <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-09-13/bigfoot.csv')

glimpse(bigfoot)
```

The Tibbles chose the Bigfoot dataset from TidyTuesday. The data originates from a publicly available database on the Bigfoot Field Researchers Organization (BFRO) website, and the dataset was created by Timothy Renner in 2017. The dataset contains 5,021 rows, with each row representing a separate Bigfoot sighting, and 28 columns, with each column providing details on the sighting.

The Tibbles opted to work with this dataset because of our shared predilection for the supernatural and the potential this information holds to prove once and for all the existence (or lack thereof) of one of the most iconic creatures in the North American imagination. It contains a wealth of observations with plenty of categorical and numerical data well-suited for comprehensive analysis and interpretation.

# Questions

# Analysis plan

A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
